On exponential sums arising from the classical doubling method

### 2025-03-03 paper.tex
Some corrections

\subsubsection{Geometric Conjugacy and Lusztig Series}
(Following chapter 13 of \cite{DigneMichel2020})
Let $\algebraicGroup{G}$ be a reductive group over $\finiteField$ and fix a torus $\algebraicGroup{T}$ defined over $\finiteField$. The \emph{dual group over $\finiteField$} of $\algebraicGroup{G}$ with respect to the torus $\algebraicGroup{T}$ corresponds to the group $\algebraicGroup{G}^*$ with a dual torus $\algebraicGroup{T}^*$ and root datum
\[
    (X^*(\algebraicGroup{T}^*),\Phi_{\algebraicGroup{T}^*},X_*(\algebraicGroup{T}^*),\Phi_{\algebraicGroup{T}^*}^\vee) \cong (X_*(\algebraicGroup{T}),\Phi_{\algebraicGroup{T}}^\vee,X^*(\algebraicGroup{T}),\Phi_{\algebraicGroup{T}})
\]
such that the isomorphism respects the Frobenius action on $X^*$ and $X_*$.

@@

A basic example of a class function which doesn't depend of geometric conjugacy 
\begin{proposition}[FIND SOMEWHERE TO CITE or put in appendix]
\label{prop:central_char}
    Any function supported on the center of $\algebraicGroup{G}(\finiteField)$ is a stable function. \calvin{maybe add in connected group} 
\end{proposition}
\begin{proof}
\calvin{todo. Here is a sketch. Use the definition of RTtheta to show for all torus theta pairs $(T,\theta)$, central element $x$ acts by scalar (in fact the scalar $\theta(x)$). This can also be showed using the character formula plus the ``only one cuspidal'' trick but it is probably overkill. Then if two pairs are conjugate, their hom dimesnion is nonzero, hence $x$ the scalar in both RTThetas} 
\end{proof}

@@

\begin{lemma}
\label{lemma:gl_invariance}
\begin{enumerate}[(i)]
	\item \label{item:gl-n-hasse-davenport-gauss-sum-implies-well-defined} In the case $G = \GL_n$ and $\rho$ is the standard representation, $g_{T,\rho}(\theta,\psi,\chi)$ depends only on the semisimple conjugacy class of $(s)$.
	\item $g_{T,\rho}(\theta,\psi,\chi)$ is well defined i.e. independent of choice of $\algebraicGroup{T}^* \subset G^*$ and $\algebraicGroup{T} \subset \GL_n$.
\end{enumerate}
\end{lemma}
\begin{proof}
\begin{enumerate}[(i)]
	\item In this case when $T = \prod_{i} \finiteFieldExtension{n_i}$ and $\theta = \prod_{i} \alpha_i$, we recover the Gauss sum
	\[
	g_{T,\rho}(\theta,\psi,\chi)  = \prod_i \tau(\alpha_i \times \chi, \fieldCharacter_{n_i}),
	\]
	which by the Hasse-Davenport relation depends only on the geometric conjugacy class of $(T,\theta)$ hence the semisimple conjugacy class of $(s_{T,\theta})$ \calvin{maybe write more or less?}. 
	\item For a Frobenius stable semisimple conjugacy class $(s)$ in $\GL_n$, we define
	\[
	g_{(s)}(\psi,\chi) := g_{T,\rho}(\theta,\psi,\chi) \textnormal{ for some } (T,\theta) \textnormal{ such that } (s_{T,\theta}) = (s)
	\]
	By (\ref{item:gl-n-hasse-davenport-gauss-sum-implies-well-defined}) this is well defined. Then
	\[
	g_{T,\rho}(\theta,\psi,\chi)  = g_{(\rho(s_{T,\theta}))}(\psi,\chi)
	\]
	which is independent of of choice of $\algebraicGroup{T}^* \subset G^*$ and $\algebraicGroup{T} \subset \GL_n$.
\end{enumerate}
\end{proof}

>>>>

\subsubsection{Geometric Conjugacy and Lusztig Series}
(Following Chapter 13 of \cite{DigneMichel2020})
Let $\algebraicGroup{G}$ be a reductive group over $\finiteField$ and fix a torus $\algebraicGroup{T}$ defined over $\finiteField$. The \emph{dual group over $\finiteField$} of $\algebraicGroup{G}$ with respect to the torus $\algebraicGroup{T}$ corresponds to the group $\algebraicGroup{G}^*$ with a dual torus $\algebraicGroup{T}^*$ and root datum
\[
    (X^*(\algebraicGroup{T}^*),\Phi_{\algebraicGroup{T}^*},X_*(\algebraicGroup{T}^*),\Phi_{\algebraicGroup{T}^*}^\vee) \cong (X_*(\algebraicGroup{T}),\Phi_{\algebraicGroup{T}}^\vee,X^*(\algebraicGroup{T}),\Phi_{\algebraicGroup{T}})
\]
such that the isomorphism respects the Frobenius action on $X^*$ and $X_*$.

@@

A basic example of a class function which doesn't depend of geometric conjugacy 
\begin{proposition}[FIND SOMEWHERE TO CITE or put in appendix]
\label{prop:central_char}
    Any function supported on the center of $\algebraicGroup{G}(\finiteField)$ is a stable function. \calvin{maybe add in connected group} 
\end{proposition}
\begin{proof}
The representation $R_T^G(\theta)$ is equal to $M \otimes_{}$
\calvin{todo. Here is a sketch. Use the definition of RTtheta to show for all torus theta pairs $(T,\theta)$, central element $x$ acts by scalar (in fact the scalar $\theta(x)$). This can also be showed using the character formula plus the ``only one cuspidal'' trick but it is probably overkill. Then if two pairs are conjugate, their hom dimesnion is nonzero, hence $x$ the scalar in both RTThetas} 
\end{proof}

@@

\begin{lemma}
\label{lemma:gl_invariance}
\begin{enumerate}[(i)]
	\item \label{item:gl-n-hasse-davenport-gauss-sum-implies-well-defined} In the case $G = \GL_n$ and $\rho$ is the standard representation, $g_{T,\rho}(\theta,\psi,\chi)$ depends only on the semisimple conjugacy class of $(s)$.
	\item $g_{T,\rho}(\theta,\psi,\chi)$ is well defined i.e. independent of choice of $\algebraicGroup{T}^* \subset G^*$ and $\algebraicGroup{T} \subset \GL_n$.
\end{enumerate}
\end{lemma}
\begin{proof}
\begin{enumerate}[(i)]
	\item In this case when $T = \prod_{i} \multiplicativegroup{\finiteFieldExtension{n_i}}$ and $\theta = \prod_{i} \alpha_i$, we recover the Gauss sum
	\[
	g_{T,\rho}(\theta,\psi,\chi)  = \prod_i \tau(\alpha_i \times \chi, \fieldCharacter_{n_i}),
	\]
	which by the Hasse-Davenport relation depends only on the geometric conjugacy class of $(T,\theta)$ hence the semisimple conjugacy class of $(s_{T,\theta})$ \calvin{maybe write more or less?}. 
	\item For a Frobenius stable semisimple conjugacy class $(s)$ in $\GL_n$, we define
	\[
	g_{(s)}(\psi,\chi) := g_{T,\rho}(\theta,\psi,\chi) \textnormal{ for some } (T,\theta) \textnormal{ such that } (s_{T,\theta}) = (s)
	\]
	By (\ref{item:gl-n-hasse-davenport-gauss-sum-implies-well-defined}) this is well defined. Then
	\[
	g_{T,\rho}(\theta,\psi,\chi)  = g_{(\rho(s_{T,\theta}))}(\psi,\chi)
	\]
	which is independent of of choice of $\algebraicGroup{T}^* \subset G^*$ and $\algebraicGroup{T} \subset \GL_n$.
\end{enumerate}
\end{proof}
